---
title: "Functions"
output: html_document
date: "2023-06-21"
---
Code gathering all handmade function used in the code package.
---

## Filtering MONTH and YEAR

Function that enables to filter a data frame of daily values on a time window split between 2 years (hydrological year for instance).
- df = dataframe of daily value of snow depth with date in YYYY-MM-DD format as the column names (df format = station*date).
- month1 = month to be filtered in year 'annee-1'
- month2 = month to be filtered in year 'annee'
- annee = year to filter out

```{r, include=FALSE}

date_filt=function(df,month1,month2,annee){

  col_year1_obs <- vector()
  col_year2_obs <- vector()

  for(i in c(1:length(colnames(df)))){
      year=as.numeric(str_split(colnames(df)[i], "-")[[1]][1])
      mois1=as.numeric(str_split(colnames(df)[i], "-")[[1]][2])
      if (year == annee & mois1 %in% month1){
       col_year1_obs[i] <- i
      } else if (year == annee+1 & mois1 %in% month2){
      col_year2_obs[i] <- i
      }
  }
  select_annee1_obs <- which(!is.na(col_year1_obs))
  select_annee2_obs <- which(!is.na(col_year2_obs))

  list_result <- list(select_annee1_obs,select_annee2_obs)
return(list_result)

}

```

## Filtering YEAR

Function that enables to filter a data frame of annual values on a selected year.
- df = dataframe of annual maxima with date in YYYY-MM-DD format as the column names (df format = station*date).
- annee = year to be filtered on

```{r, include=FALSE}

year_filt=function(df,annee){
  
  col_year <- vector()

  for(i in c(1:length(colnames(df)))){
      year=as.numeric(str_split(colnames(df)[i], "-")[[1]][1])
      if (year %in% annee){
       col_year[i] <- i
      }
  }
    select_annee <- which(!is.na(col_year))

return(select_annee)

}

```

## Intercorrelation

Function that computes a vector of one-to-one Pearson correlation values, among a group of observations.
obs = df with station*date

```{r, include=FALSE}

library(combinat)

intercor = function(obs){
  vec_corel = vector()
  nstat = dim(obs)[1]
  if (is.null(nstat)|nstat==1){
    vec_corel="single"
  }
  else if (nstat==2){
    obs1=as.numeric(obs[1,])
    obs2=as.numeric(obs[2,])
    vec_corel=cor(obs1,obs2, method = "pearson", use="pairwise.complete.obs")
  }
  
  else{combin <- combn(c(1:nstat), 2)
  
    for (i in c(1:dim(combin)[2])){
      obs1=as.numeric(obs[combin[1,i],])
      obs2=as.numeric(obs[combin[2,i],])
      vec_corel[i]=cor(obs1, obs2, method = "pearson", use="pairwise.complete.obs")
    }
  }
  
  return(vec_corel)
}

```
