---
title: "Extraction"
output: html_document
date: "2023-08-01"
---
The MTMSI reanalysis data is stored in a netcdf, with annual snow depth maxima that have been calculated on the hydrological year from Y-1/08/01 to Y/07/31. Metadata is stored aside, with gathers coordinates of the NUTS-3 barycenters and the altitudes of the timeseries. 
If there was no snow during the hydrological year of one NUTS/elevation pair, snow depth maxima is set as NA.

We extract 
- MTMSI reanalysis snow depth maxima in a dataframe with dimensions NUTS/elevation*year
- a point shapefile with NUTS-3 & elevation identification for each series, with point coordinates centered on NUTS-3 barycenter. 

---

## Library

```{r, include=FALSE}

library(plyr)
library(tidyverse)
library(tidyr)
library(dplyr)
library(ncdf4)
library(rgdal)
library(ggmap)
library(sf)

```

## Parameters

```{r, include=FALSE}

indic <- "max-sd-NS-year"

```

## Loading files

We load the NUTS3 shapefile, and netcdf of obs/mtmsi data

```{r, include=FALSE}

# MTMSI
uerra_nc <- nc_open(file=file.path("mtmsi_hist","netcdf", paste0(indic,"_UERRA.nc")))
nc_coord <- nc_open(file=file.path("mtmsi_hist", "netcdf", "MTMSI_lat_lon_ZS.nc"))
#uerra_corres <- read_csv(file.path("mtmsi_hist","LON_LAT_ZS_UERRA_EUROPE_Uerra_coord.csv"), show_col_types = FALSE)

# NUTS
shp_nut3 <- st_read(file.path("NUTS", "NUTS3_4326.shp"))

```

## Data frame with dates as colnames
We extract the annual maxima of snow depth as a dataframe.

```{r, include=FALSE}

# SNOW DEPTH
sdepth <- ncvar_get(uerra_nc, indic)
dim(sdepth)

# DATE
temps <- ncvar_get(uerra_nc, "time")
years <- lubridate::as_date(temps, origin = '1949-12-1')

# DATA FRAME 
df_date <- as.data.frame(sdepth)
colnames(df_date) <- years

save(df_date, file=file.path("mtmsi_hist", paste0(indic,"_df_date.RData")))

```

## Polyogon shapefile with NUTS3 attributes
We build a shapefile with NUTS-3 and elevation attribute of each MTMSI dataframe row. For this purpose, we extract the MTMSI coordinates from the metadata, and on the other side we compute the centroids of the NUTS-3 shapefile (build in 'nuts_extract' code). Then, we merge the information with 'st_nearest' function.

```{r, include=FALSE}

# extract data from netcdf
hist_lon <- ncvar_get(nc_coord,'LON')
hist_lat <- ncvar_get(nc_coord,'LAT')
hist_zs <- ncvar_get(nc_coord,'ZS')

# build matrix
ngrids = length(hist_lon)
hist_matrix <- matrix(0, nrow = ngrids, ncol = 3) # col: lat, lon, alt
hist_matrix[,1] <- hist_lat
hist_matrix[,2] <- hist_lon
hist_matrix[,3] <- hist_zs
    
# convert matrix into sf
hist_shp <- data.frame(hist_matrix)
colnames(hist_shp) <- c("lat","lon","alt")  
coordinates(hist_shp) <- ~lon+lat
proj4string(hist_shp) <- CRS("+init=epsg:4326") #CRS("+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0")
hist_sf <- st_as_sf(hist_shp) # we transform spatial data frame into shapefile, which is sf package friendly

# get NUTS barycenter shp for easy NUTS identification (few NUTS where no data: turkey + domtom)
shp_nuts_bary <- st_centroid(shp_nut3)

# joining the files by merging the nearest values
hist_sf_nut <- st_join(hist_sf, shp_nuts_bary, join = st_nearest_feature) # intersecting with nuts3

st_write(hist_sf_nut, file.path("mtmsi_hist", paste0(indic,"_sf_nut.shp")), delete_layer = TRUE)

```

