---
title: "MTMSI timeserie"
output: html_document
date: "2023-08-03"
---
We plot the snow depth annual maxima timeseries from one NUTS/elevation pair and its associated in-situ stations used for the evaluation (i.e. in the same NUTS region, at less than 150m elevation of difference, with at least 20 annual maxima).
---

## Library

```{r, include=FALSE}

library(plyr)
library(tidyverse)
library(tidyr)
library(dplyr)
library(sf)
library(ncdf4)
library(Metrics)
library(ggnewscale)
library(ggplot2)

```

## Parameters
The user has to specify the NUTS-3 and elevation below:

```{r, include=FALSE}

## user parameter  ##
nut<- "SE311" # USER PICK
alt=200 # USER PICK

## steady parameters
year_int = c(1962:2015)
seuil_year = 20
indic <- "max-sd-NS-year"

```

## Loading files

We load MTMSI and in situ data :
- annual maxima dataframes
- shapefile of the same dimensions for NUTS-3 and elevation selecting

```{r, include=FALSE, echo=FALSE}

# MTMSI
hist_df_date <- get(load(file.path("mtmsi_hist",paste0(indic,"_df_date.RData")))) 
hist_sf_nut <- st_read(file.path("mtmsi_hist", paste0(indic,"_sf_nut.shp"))) 

# OBS
obs_max <- get(load(file=file.path("obs", "obs_max.RData"))) 
obs_sf_nut <- st_read(file.path("obs", "obs_sf_nut.shp"))

```

## Filtering data in terms of NUTS and TIME

```{r, include=FALSE}

# MTMSI filtering
select_annee_hist <- year_filt(hist_df_date, year_int)
histrow=which(hist_sf_nut$nuts_id==nut)
hist_nuts_year = hist_df_date[histrow,select_annee_hist] # data filtered on nuts and year
j <- which(hist_sf_nut$alt[histrow]==alt) 
hist=hist_nuts_year[j,] # data filtered on nuts, year, and elevation

# OBS filtering
  # time
obs_time=obs_max[,which(colnames(obs_max)%in%year_int)]
  # nuts
obsrow <- which(obs_sf_nut$nuts_id==nut)
  # altitude
alt_temp <- obs_sf_nut$alt[obsrow]
delta_alt=abs(alt-alt_temp) # dif between selected alt and station elevations
obsrow.alt <- which(delta_alt<=150) # we keep station below 150m elevation difference
obs_time_alt=obs_time[obsrow[obsrow.alt],]

  # NA filtering
station_ok <- which(apply(obs_time_alt,1,function(x) 
            length(which(!is.na(x)))>=seuil_year)) # valid station, more than 20 missing annual maxima
nstat <- length(station_ok)
obs_ok <- obs_time_alt[station_ok,] 

# Shapefile plot-friendly dataframe
df_plot=data.frame(cbind(year_int, t(obs_ok), t(hist)))
stat_nm <- obs_sf_nut$stat_id[obsrow[obsrow.alt][station_ok]] # station id

colnames(df_plot) <- c("year",stat_nm,"UERRA") 

df_plot_shape <- df_plot %>% 
  pivot_longer(
    cols = colnames(df_plot)[2]:`UERRA`, 
    names_to = "nom",
    values_to = "value")

df_plot_shape$year <- as.numeric(df_plot_shape$year)
df_plot_shape$value <- as.numeric(df_plot_shape$value)

```

## Plot timeserie YEARLY MAX

```{r, include=FALSE}

# plotting
ggplot(df_plot_shape)+
  geom_line(data = df_plot_shape %>% filter(nom!='UERRA'), aes(y=value,x=year,colour=nom), 
            linewidth=0.5, show.legend = FALSE)+
  labs(y="Maximum snow depth (m)", x="Year")+
  scale_color_manual(name="In situ observations: ",values=c(
"#86da68","#002f51","#00e957","#c16fff","#3bc400","#d5ca36","#dd8400","#0050bd","#650ac3","#40004c","#d1ff9e",
"darkblue","grey","#b2d3ff", "#b7ffe3","#340005","#fffa95","#fc86ff","#fdb95b","#0193aa","pink",
"#00627f","#ffe494","#00333b","#e8ffdd","#8f4e00","#607100","#ffd5f7","#008e47","#ffadb8","#2c3c00",
"#146300", "#fdb95b", "#52561a", "#b2d257", "#a65630", "#cfc088"))+  # labels=TRUE
  new_scale_colour() +
  geom_line(data = (df_plot_shape %>% filter(nom=='UERRA')),aes(y=value,x=year, color="UERRA"), linewidth=0.6)+
  scale_color_manual(name='MTMSI:',values=c(UERRA="red"), labels = NULL)+
  scale_x_continuous(breaks = scales::pretty_breaks(n = 12), minor_breaks = seq(1960,2020,1))+
  #coord_cartesian(ylim=c(0,3.5))+
  #labs(title = paste0(nut, " ", alt,"m"))+
  labs(title = "(a)")+
  theme_bw()+
  theme(legend.position = "bottom", plot.title.position = 'plot', plot.title = element_text(hjust = 0.5))

ggsave(file=file.path("plot", paste0(nut,"_",alt,"tsALLstat.jpg")), dpi = 300, width = 8, height = 4)

```

