---
title: "Extract NUTS details"
output: html_document
date: "2024-06-11"
---
We extract NUTS-3 information in 3 files:
- shapefile with NUTS-3 shapes, id. Some computing is done in order to preserve the correct writing of NUTS-3 names
- 
---



## Library

```{r, include=FALSE}

library(plyr)
library(tidyverse)
library(tidyr)
library(dplyr)
library(rgdal)
library(sf)
library(readODS)
library(eurostat)

```

## Basic shapefile of NUTS-3

Basic shapefile of NUTS-3 polygons, with process to preserve NUTS-3 names

```{r, include=FALSE}

# shapefile of 1524 NUTS-3 from online download
shp_nut1524 <- st_read(file.path("NUTS", "NUTS3_ID","NUTS3_ID.shp")) %>% select(-the_geom)
shp_1524_base <- st_transform(shp_nut1524, CRS("+init=epsg:4326")) # into WGS84
  # adjust (most of) name encoding
Encoding(shp_1524_base$nuts_name) <- "UTF-8"
shp_1524_base$nuts_name <- iconv(shp_1524_base$nuts_name, from="UTF-8", to="LATIN1")

# shapefile of 1522 NUTS-3 with the right names from 'eurostat package'
euro3 <- eurostat_geodata_60_2016 %>% filter(LEVL_CODE==3)
euro3 <- euro3 %>% rename('nuts_id'="NUTS_ID") %>% select(nuts_id, NUTS_NAME)

# merging the two shapefiles data to correct 1522 nuts name (the 2 remaining are ok anyway)
shp_1524ok <- merge(shp_1524_base, st_drop_geometry(euro3), by="nuts_id", all.x=TRUE)
shp_1524f <- shp_1524ok %>% select(-nuts_name) %>% rename('nuts_name'='NUTS_NAME')
id_na <- which(is.na(shp_1524f$nuts_name))
nut_na <- shp_1524f$nuts_id[which(is.na(shp_1524f$nuts_name))]
shp_1524f$nuts_name[id_na]<-shp_1524_base$nuts_name[which(shp_1524_base$nuts_id%in%nut_na)]

# 2 shp with projection 3857 and 4326
shp_nut_3857 <- st_transform(shp_1524f, CRS("+init=epsg:3857"))
shp_nut_4326 <- st_transform(shp_1524f, CRS("+init=epsg:4326"))

st_write(shp_nut_4326, file.path("NUTS", "NUTS3_4326.shp"), delete_layer = TRUE, layer_options = "ENCODING=UTF-8")
st_write(shp_nut_3857, file.path("NUTS", "NUTS3_3857.shp"), delete_layer = TRUE, layer_options = "ENCODING=UTF-8")

```

## Shapefile of NUTS with ancillary information: relief type and ideal elevation 

Ancillary information extracted from MTMSI
- we discriminate "mountain" nuts with multiple elevation available, from "plain" with single elevation: 'relief' column is either equal to 'montagne' or to an integer which is the NUTS mean elevation of the plain NUTS
- alt_ideal: ideal mean elevation of the NUTS-3

```{r, include=FALSE}

hist_nut <- st_read(file.path("mtmsi_hist", "max-sd-NS-year_sf_nut.shp")) # dataframe with dates
shp_nut3 <- st_read(file.path("NUTS","NUTS3_4326.shp")) %>% mutate(relief=NA)
nuts_ideal <- read_ods(file.path("NUTS", "uerra5_alt_ideal_mean_nuts.ods")) # dataframe with ideal NUTS-3 mean elevation

for (n in unique(hist_nut$nuts_id)){
  
  lin_poly=which(shp_nut3$nuts_id==n)
  lin_hist=which(hist_nut$nuts_id==n)
  
  # single elevation -> keep numerical value
  if(length(lin_hist)==1){
    shp_nut3$relief[lin_poly] <- hist_nut$alt[lin_hist]
  }
  # multiple elevation -> "montagne"
  else {
    shp_nut3$relief[lin_poly] <- "montagne"
  }
}

shp_nut_detail=merge(x = shp_nut3, y = nuts_ideal %>% select(nuts_id, alt_ideal), by = "nuts_id", all.x = TRUE) %>% drop_na(relief)

st_write(shp_nut_detail, file.path("NUTS", "nuts_ideal.shp"), delete_layer = TRUE, layer_options = "ENCODING=UTF-8")

```

## Nuts average elevation with correction

We compare ideal mean elevation and MTMSI elevation bands for each NUTS-3
If the ideal mean elevation does not relate to any of the MTMSI elevation band, we put the closest one from MTMSI instead.

```{r, include=FALSE}

shp_altideal <- st_read(file.path("NUTS","nuts_ideal.shp"))
hist_nut <- st_read(file.path("mtmsi_hist", "max-sd-NS-year_sf_nut.shp"))
nuts_hist <- unique(hist_nut$nuts_id)
shp_alt_mean = shp_altideal
shp_alt_mean$alt_mtmsi=rep(NA,1517)
shp_alt_mean$id_mtmsi=rep(NA,1517)

for (n in nuts_hist){
  id_alti_ajust=which.min(abs(hist_nut$alt[which(hist_nut$nuts_id==n)]-shp_altideal$alt_ideal[which(shp_altideal$nuts_id==n)]))
  shp_alt_mean$alt_mtmsi[which(shp_alt_mean$nuts_id==n)] <- hist_nut$alt[which(hist_nut$nuts_id==n)][id_alti_ajust]
  shp_alt_mean$id_mtmsi[which(shp_alt_mean$nuts_id==n)] <- which(hist_nut$nuts_id==n)[id_alti_ajust]
}

shp_alt_mean3857 <- st_transform(shp_alt_mean, 3857)

st_write(shp_alt_mean3857, file.path("NUTS", "shp_altMean3857.shp"), append=FALSE, layer_options = "ENCODING=UTF-8")
st_write(shp_alt_mean, file.path("NUTS", "shp_altMean4326.shp"), append=FALSE, layer_options = "ENCODING=UTF-8")

```
