---
title: "Merging observation extracts"
output: html_document
date: "2024-06-07"
---
We merge the in-situ data sets of annual maxima: European Alp, Germany, Finland, Ecad.

One paragraph is devoted to manually checked some stations, either both at the same location or with the same name.
For data sets that needed to be cropped (alps german and ecad ones), two files exist: 'xxx_FULL' (full one) and 'xxx' (cropped one), both for the annual maxima and the shapefile. Indeed, annual maxima data frames and shapefiles require the same dimension in order to relate each serie to its station.
---

## Library

```{r, include=FALSE}

library(plyr)
library(tidyverse)
library(tidyr)
library(dplyr)
library(ncdf4)
library(rgdal)
library(sf)
library(readODS)

```

## Loading files

```{r, include=FALSE}

# annual maxima
ger_max <- get(load(file=file.path("obs", "Allemagne", "df_ger_year.RData")))
alp_max <- get(load(file=file.path("obs", "Alpes", "df_alp_year.RData")))
fin_max <- get(load(file.path("obs", "Finland", "df_fin_year.RData")))
eur_max <- get(load(file.path("obs", "Ecad", "df_ecad_year.RData"))) # after intersection
  
# shapefiles
ger_sf <- st_read(file.path("obs", "Allemagne", "ger_sf_nut.shp"))
alp_sf <- st_read(file.path("obs", "Alpes", "alp_sf_nut.shp"))
fin_sf <- st_read(file.path("obs", "Finland", "fin_sf_nut.shp"))
eur_sf <- st_read(file.path("obs", "Ecad", "ecad_sf.shp")) 

# NUTS3
shp_nut3 <- st_read(file.path("NUTS","NUTS3_3857.shp"))

```

## Data frame with dates as colnames

We bind rows of annual maximum files.

```{r, include=FALSE}

# yearly max of alp-ger-fin 
obs_max=rbind.fill(alp_max, ger_max, fin_max)
save(obs_max, file=file.path("obs", "obs_maxAGF.RData"))

# yearly max of alp-ger-fin-ecad
obs_max=rbind.fill(alp_max, ger_max, fin_max, eur_max)
save(obs_max, file=file.path("obs", "obs_max.RData"))

```

## Point shapefile with NUTS3 attribute
We bind rows of shapefiles.

```{r, include=FALSE}

obs_sf_nut <- rbind(alp_sf, ger_sf, fin_sf %>% select(-stat_num))
    
st_write(obs_sf_nut, file.path("obs", "obs_sf_nutAGF.shp"), delete_layer = TRUE)

# shp max of alp-ger-fin-ecad set
obs_sf_nut <- rbind(alp_sf, ger_sf, fin_sf %>% select(-stat_num), eur_sf %>% select(-stat_id) %>% rename('stat_id'=stat_name))
    
st_write(obs_sf_nut, file.path("obs", "obs_sf_nut.shp"), delete_layer = TRUE)

```

## REMOVING DUPLICATES FROM EACH DATA SET ##
We look for duplicate stations:
- within Alp set (manual check within German set showed no duplicate)
- between Alp and German set 
- between Ecad and Alp-German-Finland set

```{r, include=FALSE, echo=FALSE}

#-------------------------------------------------------------
#### CHECKING ALPS ####
# We look for stations with identical location, by buidling a 
# buffer of the shapefile.
#-------------------------------------------------------------

alp_sf <- st_read(file.path("obs", "Alpes", "alp_sf_nutFULL.shp"))
alp_maxFULL <- get(load(file=file.path("obs", "Alpes", "df_alp_yearFULL.RData")))

# build buffer to compute intersections
# alp_buff200 <- st_buffer(alp_sf, dist=200)
# st_write(alp_buff200, file.path("obs", "Alpes", "alpBUFF200.shp"), delete_layer = TRUE)
# 
# alp_check_identical <- st_contains(alp_sf, alp_sf)
# alp_check_buf <- st_contains(alp_buff200, alp_buff200)
# #
# # # df summarizing intersections
# df_stat = matrix(ncol=3)
# for (i in c(1:2893)){
#   vect_stat <- rep(NA,3)
#   stat_with <- alp_sf$stat_id[alp_check_buf[[i]]]
#     if(length(stat_with)>=2){
#       id_fill <- which(!is.na(stat_with))
#       vect_stat[id_fill]<-stat_with
#       df_stat <- rbind(df_stat, t(as.matrix(vect_stat)))
#     }
# }

# conclusion: removing stations and saving file
 remove_statname <- c('Zernez_CH_METEOSWISS', 'Zervreila_CH_METEOSWISS', 'Val_d_isere_jos', 'Val_d_isere', 'Ulrichen_CH_METEOSWISS', 'Tignes_sapc', 'St_christophe_38375400', 'St_Moritz_CH_METEOSWISS', 'Maljasset', 'Scuol_CH_METEOSWISS', 'Schwagalp_CH_METEOSWISS', 'San_Bernardino', 'Punt_la_Drossa', 'Pelvoux_st_antoine', 'Murren_CH_METEOSWISS', 'Passo_del_Bernina', 'Megeve_auxi', 'Les_portes_en_valgaudemar', 'Le_tour', 'Diga_Goillet', 'Diga_Cignanaz', 'Isola', 'Innerferrera_CH_METEOSWISS', 'Grimsel_Hospiz_CH_METEOSWISS', 'Gsteig_Gstaad', 'Gadmen_CH_METEOSWISS', 'Fionnay_CH_METEOSWISS', 'Ceillac', 'Braunwald_CH_METEOSWISS', 'Bourg_St_Pierre_CH_METEOSWISS', 'Bessans', 'Adelboden_CH_METEOSWISS')

stat_row <- which(alp_sf$stat_id %in%remove_statname)

# # filtering out
alp_sfCROP <- alp_sf[-stat_row,]
alp_maxCROP <- alp_maxFULL[-stat_row,] # stations timeseries are also removed

st_write(alp_sfCROP, file.path("obs", "Alpes", "alp_sf_nut.shp"), delete_layer = TRUE)
save(alp_maxCROP, file = file.path("obs", "Alpes", paste0("df_alp_year.RData")))

# # #-------------------------------------------------------------
# # #### checking GERMAN vs ALP SET ####
# In a similar way to above, we look for duplicate stations by
# building a buffer around one shapefile.
# # #-------------------------------------------------------------

ger_sf_full <- st_read(file.path("obs", "Allemagne", "ger_sf_nutFULL.shp"))
alp_sf <- st_read(file.path("obs", "Alpes", "alp_sf_nut.shp"))

alp_buff200 <- st_buffer(alp_sf, dist=200)
contain_list <- st_contains(alp_buff200, ger_sf_full)
german_dup <- unlist(contain_list)

# filter out shapefile
ger_crop <- ger_sf_full[-german_dup,]
st_write(ger_crop, file.path("obs", "Allemagne", "ger_sf_nut.shp"), delete_layer = TRUE)

# filter out yearly max
ger_max <- get(load(file.path("obs", "Allemagne", "df_ger_yearFULL.RData")))
ger_maxcrop <- ger_max[-german_dup,]
save(ger_maxcrop, file=file.path("obs", "Allemagne", "df_ger_year.RData"))

# # #-------------------------------------------------------------
# # #### CHECKING ECAD vs ALPGERFIN SET ####
# We have merged alps - german - finland set in order to remove 
# the stations from ecad that were already in the set.
# # #-------------------------------------------------------------

# # Alps data
AGF_sf <- st_read(file.path("obs", "obs_sf_nutAGF.shp"))
# # buidling buffer at 200m
AGF_buf200 <- st_buffer(AGF_sf, dist=200)

# # EcAD data
ecad_sfFULL <- st_read(file.path("obs", "Ecad", "ecad_sf_nut_full.shp"))
ecad_yearFULL <- get(load(file.path("obs", "Ecad", "eur_yearFULL.RData")))
# 
# # intersection
intersect <- st_contains(AGF_buf200, ecad_sfFULL)
# 
df_stat = matrix(ncol=3)
for (i in c(1:7824)){
  vect_stat <- rep(NA,3)
  stat_with <- ecad_sfFULL$stat_name[intersect[[i]]]
    if(length(stat_with)>=1){ ### two data sets, so only one station interesected!
      id_fill <- which(!is.na(stat_with))
      vect_stat[id_fill]<-stat_with
      df_stat <- rbind(df_stat, t(as.matrix(vect_stat)))
    }
}
# 
# # removing stations from ecad: WARNING some
stat1 <- unique(as.vector(df_stat[which(!is.na(df_stat[,1])),1]))
stat2 <- unique(as.vector(df_stat[which(!is.na(df_stat[,2])),2]))
stat_row <- which(ecad_sfFULL$stat_name %in% stat1)
# 
new_ecadsf <- ecad_sfFULL[-stat_row, ]
new_ecadyear <- ecad_yearFULL[-stat_row, ]

st_write(new_ecadsf, file.path("obs", "Ecad", "ecad_sf.shp"), delete_layer = TRUE)
save(new_ecadyear, file=file.path("obs", "Ecad", "df_ecad_year.RData"))

```


