---
title: "Extracting in-situ data from the European Alps"
output: html_document
date: "2023-06-14"
---
The initial netcdf contains daily values of snow depth ('snd' variable, in cm) from 1961.01.01 to 2020.12.31.

(1) We extract those daily values. Then we compute the annual maxima for each station, on the hydrological year from Y-1/08/01 to Y/07/31, for stations with less than 10% missing values on 'likely maxima period' from Y-1/11/01 to Y/04/30.

(2) We compute a shapefile of the stations, using the coordinates and station names from the netcdf file, and the NUTS-3 shapefile attributes.

---

## Library

```{r, include=FALSE}

library(plyr)
library(tidyverse)
library(tidyr)
library(dplyr)
library(ncdf4)
library(rgdal)
library(sf)
library(readODS)

```

## Parameters

```{r, include=FALSE}

seuil_saison=0.1 
year_int=c(1962:2015) 

```

## Loading files

```{r, include=FALSE}

# obs
ncd_alp <- nc_open(file=file.path("obs", "Alpes", "snd_Alps_1961_2020.nc"))

# station id
alp_name=as.character(ncd_alp$dim$Station_Name$vals)

# NUTS3
shp_nut3 <- st_read(file.path("NUTS","NUTS3_4326.shp"))

```

## Alpine: From daily to yearly max 

- We compute NA rate for each station on winter season (November-April)
- We then compute their max on full year (August to July)
- We remove timeseries full of NA

```{r, include=FALSE}

# alps: monthly data
sdepth <- ncvar_get(ncd_alp,'snd')/100
temps <- ncvar_get(ncd_alp, "time")
jour <- lubridate::as_date(temps, origin = '1961-01-01')

# alps df dim station*days
df_alps_day <- as.data.frame(sdepth)
colnames(df_alps_day) <- jour
save(df_alps_day, file=file.path("obs", "Alpes", "df_alps_day.RData"))

obs_df_date <- get(load(file.path("obs", "Alpes", "df_alps_day.RData")))
obs_data_year <- list()

### Loop on each year ###
i=0
for (y in year_int){

  i=i+1

  ################## FILTERING ON NA RATE ON WINTER SEASON (NOVEMBER->APRIL) ##################

    # filtering winter season
  sel_annee <- date_filt(obs_df_date, month1=c(11:12), month2 = c(1:4), annee = y-1) 
  saison_na <- c(sel_annee[[1]],sel_annee[[2]])

    # we save indices of stations with NA exceeding threshold
  tx_na <- apply(obs_df_date[,saison_na], 1, function(x) length(which(is.na(x)))/length(saison_na)) # vector of na amount per station
  tx_na[which(tx_na>=seuil_saison)] <- NA # values above thresholds are set as NA

  ################## COMPUTING MAX ON FULL YEAR (AUGUST year-1 -> JULY year+1) ##################

    # filtering year centered on winter y-1 > y
  sel_annee2 <- date_filt(obs_df_date, month1=c(8:12), month2 = c(1:7), annee = y-1)
  saison_max <- c(sel_annee2[[1]],sel_annee2[[2]])

    # allocating NA to timeseries full of NA or with NA rate too high
  obs_max <- rep(NA,dim(obs_df_date)[1]) # initialyzing vector of obs max
  finite_value <- which(apply(obs_df_date[,saison_max],1,function(x) length(which(is.na(x)))<length(x))) # indices of series with at least one value
  obs_max[finite_value] <- apply(obs_df_date[finite_value,saison_max],1,max,na.rm=TRUE) # replacing original NA with max where possible (at least one value)
  obs_max[is.na(tx_na)]<-NA # replacing with NA where NA rate was above threshold 0.1

  obs_data_year[[i]]<-obs_max

}

alp_max <-  as.data.frame(do.call(cbind, obs_data_year))
colnames(alp_max) <- year_int
save(alp_max, file = file.path("obs", "Alpes", paste0("df_alp_yearFULL.RData")))

```

## Point shapefile with NUTS3 attribute

```{r, include=FALSE}

  # extract coordinate data from netcdf
alp_lon <- ncvar_get(ncd_alp,'lon')
alp_lat <- ncvar_get(ncd_alp,'lat')
alp_ZS <- ncvar_get(ncd_alp,'ZS')
  
  # build matrix
ngrids = length(alp_lat)
alp_matrix <- matrix(0, nrow = ngrids, ncol = 3) 
alp_matrix[,1] <- alp_lat
alp_matrix[,2] <- alp_lon
alp_matrix[,3] <- alp_ZS
      
  # convert matrix into sf
alp_shp <- data.frame(alp_matrix)
colnames(alp_shp) <- c("lat","lon","alt")
coordinates(alp_shp) <- ~lon+lat
proj4string(alp_shp) <- CRS("+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0")
alp_sf <- st_as_sf(alp_shp) # we transform spatial data frame into shapefile, which is plotting friendly
alp_sf_nut <- st_join(alp_sf, shp_nut3, join = st_nearest_feature) # joining with nut3 id
alp_sf_nut$stat_id <- alp_name
  
st_write(alp_sf_nut, file.path("obs", "Alpes", "alp_sf_nutFULL.shp"), delete_layer = TRUE)
  
```

