---
title: "mtmsi_rmse"
output: html_document
date: "2023-07-13"
---

We compute list of scores (rmse, mae, bias, KGE, correlation, station median intercorrelation) related altitude, for each NUTS/elevation pair and the number of stations involved in the score.

---

## Library

```{r, include=FALSE}

library(plyr)
library(tidyverse)
library(tidyr)
library(dplyr)
library(ncdf4)
library(rgdal)
library(sf)
library(Metrics)
library(hydroGOF)

```

## Parameters

```{r, include=FALSE}

seuil_year= 20 
year_int=c(1962:2015)
indic <- "max-sd-NS-year"

```

## Loading files

We load 2 types of files:
- obs_maxfull/ hist_date = annual maxima with date as column names
- obs_sf_nut/ hist_nut = shapefile with corresponding nut, and elevation 

```{r, include=FALSE}

# OBS
obs_maxfull <- get(load(file = file.path("obs","obs_max.RData")))
obs_maxfull <- mutate_all(obs_maxfull, function(x) as.numeric(as.character(x)))
obs_sf_nut <- st_read(file.path("obs", "obs_sf_nut.shp"))

# weird stations
obs_maxfull[12786,] <- NA # only 0 and NA causing score error
obs_maxfull[which(obs_sf_nut$nuts_id=='HR034'),] <- NA # very high values at 300m
obs_maxfull[which(obs_sf_nut$nuts_id=='HR037'),] <- NA # very high values at 200m
#obs_maxfull[which(obs_sf_nut$nuts_id=='HR04B'),] <- NA # very high values at 100m

# MTMSI
hist_nut <- st_read(file.path("mtmsi_hist",paste0(indic,"_sf_nut.shp"))) # dataframe with nuts correspondances
hist_date <- get(load(file.path("mtmsi_hist", paste0(indic,"_df_date.RData")))) # dataframe with dates
# NUTS ID
nuts_hist <- unique(hist_nut$nuts_id)
# MTMSI elevations
alt_hist <- hist_nut$alt

```

## Filtering on time window

```{r, include=FALSE}

# Filtering MTMSI 
select_annee_hist <- year_filt(hist_date, year_int)
hist_datafilt = hist_date[,select_annee_hist]

# Filtering obs 
select_annee_obs <- year_filt(obs_maxfull, year_int)
obs_max = obs_maxfull[,select_annee_obs]

```

## Computing scores per NUTS-3 and altitude

We compute scores, comparing UERRA and ground truth data, following multiple conditions:
- presence of station
- presence at similar altitude
- more than 20 annual maxima available

```{r, include=FALSE}

rmse_list<-list()
bias_list<-list()
mae_list<-list()
cor_list<-list()
altitude_list <- list()
station_list <- list()
intercor_min_l <- list()
intercor_med_l <- list()
kge_list <- list()
stat_row_l <- list()

for (i in c(1:length(nuts_hist))){ # loop on mtmsi nuts3 
  
  obs_row <- which(obs_sf_nut$nuts_id==nuts_hist[i]) # station within the selected nut3
  
  rmse <- list()
  biais <- list()
  mae <- list()
  corr <- list()
  alt_vec <- vector()
  stat_vec <- vector()
  intercor_min <- vector()
  intercor_med <- vector()
  kkge <- list()
  stat_row <- list()

  if (length(obs_row)>=1){ # if there is at least one station in the nut
    
    mtmsi_rows <- which(hist_nut$nuts_id==nuts_hist[i]) # mtmsi filtered on selected nuts3
    altitudes <- alt_hist[mtmsi_rows] # mtmsi altitude available for the nuts3
    
    for (j in c(1:length(altitudes))){ # loop on altitudes 
      
      alt_temp <- altitudes[j]
      delta_alt=abs(obs_sf_nut$alt[obs_row]-alt_temp) # dif between selected alt and station elevations
      select_altitude <- which(delta_alt<=150) # we keep station within 150m elevation delta
      
      if(length(select_altitude)>=1){ # if there is at least one station at same altitude
        
        obs_row_150 <- obs_row[select_altitude] # selected station based on nuts and altitude selection
        obs_max_alt <- obs_max[obs_row_150,] # resulting data frame of max
        annee_ok <- apply(obs_max_alt, 1, function(x) length(which(!is.na(x)))) # number of annual maxima available for every station
        obs_nafilt <- obs_max_alt[which(annee_ok>=seuil_year),] # removing data from station with less than  20 annual maxima

        mtmsi_row_alt <- mtmsi_rows[j] # mtmsi row of selected nut and altitude
        int=as.numeric(hist_datafilt[mtmsi_row_alt,]) # extracting timeseries of that row
        df_hist <- data.frame(year=year_int,value=int) # saving in a df
        
        if(max(annee_ok)>=seuil_year){ # if at least one station has enough annual max
          
          rmse[[j]] <- apply(obs_nafilt, 1, function(x) 
            rmse(x[which(!is.na(x))], df_hist$value[which(!is.na(x))])/mean(x, na.rm = TRUE))
          biais[[j]] <- apply(obs_nafilt, 1, function(x) 
            -bias(x[which(!is.na(x))], df_hist$value[which(!is.na(x))]))
          mae[[j]] <- apply(obs_nafilt, 1, function(x) 
            mae(x[which(!is.na(x))], df_hist$value[which(!is.na(x))])/mean(x, na.rm = TRUE))
          corr[[j]] <- apply(obs_nafilt, 1, function(x) 
            cor(x[which(!is.na(x))], df_hist$value[which(!is.na(x))], method = "pearson"))
          kkge[[j]] <- apply(obs_nafilt, 1, function(x) 
            KGE(x[which(!is.na(x))], df_hist$value[which(!is.na(x))], method = "2012")) 
          
          stat_row[[j]] <- obs_row_150[which(annee_ok>=seuil_year)] # rows of selected station, to keep track of which station were used
          
          alt_vec[j]<-altitudes[j] # altitude evaluated
          
          cor_interstat=intercor(obs_nafilt) # correlation between station
          
          if (length(which(!is.na(cor_interstat)))>=1){ # if there is more than 1 station, we summarize the intercor score
           intercor_min[j]=min(cor_interstat, na.rm=TRUE)
           intercor_med[j]=median(cor_interstat, na.rm=TRUE)
          }
          else { # if there only one station
           intercor_min[j]=NA
           intercor_med[j]=NA
          }
          
          stat_vec[j] <- nrow(obs_nafilt) # number of station used = with enough data

        } else {        # if there is not enough annual maxima from obs
          rmse[[j]] <- NA
          biais[[j]] <- NA
          mae[[j]] <- NA
          corr[[j]] <- NA
          kkge[[j]] <- NA
          alt_vec[j] <- altitudes[j]
          stat_vec[j] <- NA
          intercor_min[j] <- NA
          intercor_med[j] <- NA
          stat_row[[j]] <- NA
        }
      
      } else {        # if there is no station elevation close enough
          rmse[[j]] <- NA
          biais[[j]] <- NA
          mae[[j]] <- NA
          corr[[j]] <- NA
          kkge[[j]] <- NA 
          alt_vec[j] <- altitudes[j]
          stat_vec[j] <- NA
          intercor_min[j] <- NA
          intercor_med[j] <- NA
          stat_row[[j]] <- NA
      }

    } 
    
   rmse_list[[i]]<-rmse
   bias_list[[i]]<-biais
   mae_list[[i]]<-mae
   cor_list[[i]]<-corr
   kge_list[[i]]<-kkge 
   altitude_list[[i]]<-alt_vec
   station_list[[i]] <- stat_vec
   intercor_min_l[[i]] <- intercor_min
   intercor_med_l[[i]] <- intercor_med
   stat_row_l[[i]] <- stat_row
   
    }else {        # if there is not any station in the nut
      rmse_list[[i]]<-NA
      bias_list[[i]]<-NA
      mae_list[[i]]<-NA
      cor_list[[i]]<-NA
      kge_list[[i]]<-NA 
      altitude_list[[i]] <- NA
      station_list[[i]] <- NA
      intercor_min_l[[i]] <- NA
      intercor_med_l[[i]] <- NA
      stat_row_l[[i]] <- NA
    }
      
}

save(rmse_list, file=file.path("mtmsi_hist", "score", "rmse_list.RData"))
save(bias_list, file=file.path("mtmsi_hist", "score", "bias_list.RData"))
save(mae_list, file=file.path("mtmsi_hist", "score", "mae_list.RData"))
save(cor_list, file=file.path("mtmsi_hist", "score", "cor_list.RData"))
save(altitude_list, file=file.path("mtmsi_hist", "score", "alt_list.RData"))
save(station_list, file=file.path("mtmsi_hist", "score", "stat_list.RData"))
save(intercor_min_l, file=file.path("mtmsi_hist", "score", "intcormin_list.RData"))
save(intercor_med_l, file=file.path("mtmsi_hist", "score", "intcormed_list.RData"))
save(kge_list, file=file.path("mtmsi_hist", "score", "kge_list.RData"))
save(stat_row_l, file=file.path("mtmsi_hist", "score", "statrow_list.RData"))

```


