---
title: "Shapefile of score on NUTS-3 mean elevation"
output: html_document
date: "2024-06-13"
---
We compute a shapefile of the selected score, on the mean elevation of all NUTS-3 (if available)
---
## Library

```{r, include=FALSE}

library(tidyverse)
library(tidyr)
library(rgdal)
library(sf)
library(ggmap)
library(ggplot2)
library(plyr)
library(dplyr)
library(sf)

```

## Parameters

```{r, include=FALSE}

sel.score='cor' # one of: bias cor kge
indic <- "max-sd-NS-year"

```

## Loading files

```{r, include=FALSE}

# MTMSI details
hist_nut <- st_read(file.path("mtmsi_hist", paste0(indic,"_sf_nut.shp"))) # shapefile with NUTS-3 details
nuts_hist <- unique(hist_nut$nuts_id)

# altitude and stations
altitude_list <- get(load(file.path("mtmsi_hist","score", "alt_list.RData"))) # evaluated altitude for each nuts
stat_list <- get(load(file.path("mtmsi_hist","score", "stat_list.RData"))) # amount of stations for each NUTS/elev evaluation

# SCORE
score_list <- get(load(file.path("mtmsi_hist","score",paste0(sel.score,"_list.RData")))) # score values for each NUTS

# NUT
shp_nut3 <- st_read(file.path("NUTS","NUTS3_4326.shp")) # nuts shapefile
shp_nut_elev <- st_read(file.path("NUTS","shp_altMean4326.shp")) # nuts with mean elevation shapefile

```

## Data frame of score and station count per NUT3 unit

```{r, include=FALSE}

# extracting list of scores into dataframe
df_score <- data.frame()

for (i in c(1:length(nuts_hist))){
  val=unlist(lapply(score_list[[i]], function(x) median(unlist(x), na.rm=TRUE))) # summarizing score values for each NUTS/elevation: median of values generated from the pool of stations
  alt=unlist(altitude_list[[i]]) # altitude assessed
  stat=unlist(stat_list[[i]]) # number of stations involved
  df_temp=data.frame(valeur=val, altitude=alt, station=stat, nuts_id=nuts_hist[i])
  df_score=rbind(df_score,df_temp)
} 

```

## Adding mean elevation 
We intersect the score shapefile generated above (where 1 row = 1 NUTS/elev) with the NUTS-3 mean elevation shapefile (where 1 row = 1 NUTS) 

```{r, include=FALSE}

df_score2 = data.frame()

for (n in nuts_hist){
  temp_alt = shp_nut_elev$alt_mtmsi[which(shp_nut_elev$nuts_id==n)]
  df_temp = df_score %>% filter(nuts_id==n) %>% filter(altitude==temp_alt)
  
  df_score2 = rbind(df_score2,df_temp)
}

df_summary=df_score2 %>% na.omit()

```

## Shapefile of score and nb_stat
In order to have all NUTS-3 details, we merge the above shapefile with the one of NUTS-3

```{r, include=FALSE}

shp_score <- merge(shp_nut3,df_summary,by='nuts_id')

st_write(shp_score, file.path("carte", "data", paste0(sel.score,"_all.shp")), append=FALSE)

```

